---
title: "Data Visualization"
author: "Quan Gan, Yuxuan chen"
date: "11/3/2021"
output: 
  html_document:
    toc: true
---

# Data Visualization

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
# load libraries
library(tidyverse)
library(knitr)
library(ggthemes)
```

```{r}
# load data in
LCA_application <- read_csv("Data/LCA_cleaned.csv")
```
## Question 1

> Question: What companies have high expectations of international workers arcoss years?

In the data set, the employer_name variable is the name of company. We need to
figure what companies have more LCA application from 2016 to 2020.


```{r}
# Find the top 10 companies that have most LCA applications.
LCA_application %>% 
  select(employer_name) %>% 
  add_count(employer_name) %>% 
  arrange(desc(n)) %>% 
  unique() %>% 
  top_n(10) -> top_10_company
  
```

```{r}
# Count the LCA applications of top 10 companies across years
LCA_application %>% 
  filter(employer_name %in% top_10_company$employer_name) %>% 
  select(employer_name, year) %>% 
  group_by(year) %>% 
  add_count(employer_name) %>% 
  arrange(year, desc(n)) %>% 
  unique() -> company_application_number 
```

Visualize the top 5 companies across years

```{r}
company_application_number %>% 
  group_by(employer_name, year) %>% 
  ggplot(aes(x = year,
             y = n,
             fill = year)) +
  geom_col(position = "dodge") + 
  facet_wrap(~employer_name, ncol = 3) +
  coord_flip() +
  labs(y = "Number of LCA applications") +
  scale_color_colorblind() 

```

This plot shows the top 10 companies in total that apply for LCA applications from 2016 to 2020.
Each subplot shows their applications trends. We could know for Google LLC and cognizant technology solutions us corp,
they don't have LCA applications in 2016 and 2017. Half of top 10 companies increase their
LCA applications. In addition, I choose the top 10 companies that have most LCA 
applications in total rather than choosing each year top 10 companies because some 
companies might only desire foreign workers in a specific year and analyzing overall top 10 
companies could illustrate the recruiting trends.

## Question 2

> Question: Would the prevailing wage rate and paid wage rate influence the approval of LCA application?

In LCA, the prevailing wage is defined as "the average wage paid to similarly employed
workers in the requested occupation in the area of intended employment" from 
https://www.usavisanow.com/h-1b-visa/h1b-visa-resources/prevailing-wage/.

The paid wage rate is the actually wage paied to the foreign employee. 


```{r}
# count case_status
LCA_application %>% count(case_status) %>% 
  kable()
```

In this data set, there are over 300 million records but only 1% applications failed,
which is extremely unbalanced. In order to explore the correlation among prevailing 
wage rate, paid wage rate, and the approval of LCA application, we decide to sample
same amount record from data set.

```{r}

# set seed
set.seed(1)

# sample records from dataset
certified <- LCA_application %>% 
  filter(case_status == "certified") %>% 
  sample_n(10000)

denied <- LCA_application %>% 
  filter(case_status == "denied") %>% 
  sample_n(10000)


LCA_sample <- rbind(certified, denied)

# The wage_rate_from_year is the minimum paid wage rate and wage_rate_to_year is the
# maximum paid wage rate. So, we need to calculate the mean wage rate.
LCA_sample <- LCA_sample %>% 
  mutate(mean_wage_rate_year = ifelse(wage_rate_to_year == 0, 
                                      wage_rate_from_year,
                                      (wage_rate_from_year + wage_rate_to_year)/2))

LCA_sample <- LCA_sample %>% 
  mutate(difference_prevailing_wage = mean_wage_rate_year - prevailing_wage_year)

```


```{r}
LCA_sample %>% 
  group_by(case_status) %>% 
  ggplot(aes(x = factor(year),
             y = prevailing_wage_year,
             fill = case_status)) +
  geom_boxplot() +
  coord_cartesian(ylim=c(0,150000)) +
  labs(x = "Year",
       y = "Prevailing wage rate unit year",
       fill = "Case Status") +
  scale_color_colorblind() 

  

```

```{r}
LCA_sample %>% 
  group_by(case_status) %>% 
  ggplot(aes(x = factor(year),
             y = mean_wage_rate_year,
             fill = case_status)) +
  geom_boxplot() +
  coord_cartesian(ylim=c(0,150000))+
  labs(x = "Year",
       y = "Paid wage rate unit year",
       fill = "Case Status")
```
```{r}
LCA_sample %>% 
  group_by(case_status) %>% 
  ggplot(aes(x = factor(year),
             y = difference_prevailing_wage,
             fill = case_status)) +
  geom_boxplot() +
  coord_cartesian(ylim=c(-10000,25000))+
  labs(x = "Year",
       y = "Difference of prevailing wage rate and paid wage rate",
       fill = "Case Status")
```

The means of prevailing wage rate and paid wage rate of refused LCA applications 
are lower than the means of certified LCA applications, as seen in the first two box plots. 
The mean gap between paid and prevailing pay rates is greater than the mean difference 
between paid and prevailing wage rates.
The third box plots shows that the applicants receive the denied results because 
the employers don't pay enough wage rate comparing to the prevailing wage rate.
In addition, some paid wage rates of denied applications are less than the 
prevailing wage rate but certified applications always provide wage rate larger 
than prevailing wage. In the data, there are some extreme high wage rates, which 
are noise. In order to help the users understand the visualization better, I change 
the y-axis so the audiences could focus on the mean values.

## Question 3

> Question: What is the trends of data science related jobs?

In this question, we will analyze four popular data science related jobs, machine learning,
data scientist, data engineer, and data analyst. How would 
their demand change? How would their wage rate change?

```{r}
# filter original data set and select data science related jobs.
ds_occupation <- c("machine learning", "data scientist", "data engineer", "data analyst")

LCA_application <- LCA_application %>% 
  mutate(occupation = case_when(grepl("machine learning", job_title, fixed = TRUE) ~ "machine learning",
                                grepl("data scientist", job_title, fixed = TRUE) ~ "data scientist",
                                grepl("data engineer", job_title, fixed = TRUE) ~ "data engineer",
                                grepl("data analyst", job_title, fixed = TRUE) ~ "data analyst",
                                TRUE ~ job_title))

LCA_application %>% 
  filter(occupation %in% ds_occupation) %>% 
  count(occupation) %>% 
  kable()
```


```{r}
# Visualizing LCA applications amount of data science jobs.
LCA_application %>% 
  filter(occupation %in% ds_occupation) %>% 
  group_by(occupation) %>% 
  ggplot(aes(x = year, fill = factor(year))) +
  geom_bar(stat="count", position = "dodge") +
  geom_label(stat='count', aes(label=..count..), hjust = -0.1) +
  scale_y_continuous(limits=c(0, 6000)) +
  scale_color_colorblind() +
  coord_flip() +
  facet_wrap(~occupation) + 
  labs(x = "Year",
       y = "Number of Applications",
       fill = "Year") 
```
This bar chart depicts the number of LCA applications. of data science-related 
professions over time. With each passing year, the need for data science-related 
occupations grows. Particularly, the data engineer occupation growth fast. 
In addition, to assist viewers comprehend the graphic, I indicate the specific 
number of LCA applications. Despite the fact that data science is becoming more 
popular, it still still accounts for a small portion of demand in the US market, 
which totals 300 million applications.


```{r}
# Visualizing the prevailing wage rage of data science job
LCA_application %>% 
  filter(occupation %in% ds_occupation) %>% 
  group_by(occupation, year) %>% 
  summarise(mean_wage = mean(prevailing_wage_year)) %>% 
  ggplot(aes(x = year, y = mean_wage, color = occupation)) +
  geom_line(size = 1) +
  labs(x = "Year",
       y = "Mean of Prevailing wage Rate",
       fill = "Occupation")
```
```{r}
# Visualizing the mean wage rage of data science job
LCA_application %>% 
  filter(occupation %in% ds_occupation) %>% 
  group_by(occupation, year) %>% 
  summarise(mean_wage = mean(wage_rate_from_year)) %>% 
  ggplot(aes(x = year, y = mean_wage, color = occupation)) +
  geom_line(size = 1) +
  labs(x = "Year",
       y = "Mean of Minimum paid wage Rate",
       fill = "Occupation")
```

The two graphs above show pay trends for data science positions. Except for the 
data engineer occupation, which has seen a significant increase in wage rates in 2018, 
other data science jobs have seen just minor changes. They raise the paid minimum 
wage rate for those who seek to recruit overseas labor. The data scientist occupation 
has the quickest income increase and is also one of the most in-demand positions 
in the data science field.We use different color lines to represent different jobs 
and alter the line width to assist users comprehend the graphic.

