---
title: "Data modeling"
author: "Quan Gan, Yuxuan chen"
date: "11/16/2021"
output: html_document
---

# Data Modeling

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
# load libraries
library(tidyverse)
library(knitr)
library(effects)
library(ggthemes)
```

```{r}
# load data in
LCA_application <- read_csv("Data/LCA_cleaned.csv")
```

In this data set, there are over 300 million records but only 1% applications failed,
which is extremely unbalanced. In order to ensure the fairness, we sample approval 
and denied cases in the same size and use the new subset of LCA applications to train 
the models.

```{r}

# set seed
set.seed(1)

# sample records from dataset
certified <- LCA_application %>% 
  filter(case_status == "certified") %>% 
  sample_n(10000)

denied <- LCA_application %>% 
  filter(case_status == "denied") %>% 
  sample_n(10000)


LCA_sample <- rbind(certified, denied)

# The wage_rate_from_year is the minimum paid wage rate and wage_rate_to_year is the
# maximum paid wage rate. So, we need to calculate the mean wage rate.
LCA_sample <- LCA_sample %>% 
  mutate(mean_wage_rate_year = ifelse(wage_rate_to_year == 0, 
                                      wage_rate_from_year,
                                      (wage_rate_from_year + wage_rate_to_year)/2))

LCA_sample <- LCA_sample %>% 
  mutate(difference_prevailing_wage = mean_wage_rate_year - prevailing_wage_year)

```


## Model 1

> Question: What is the relationship of year and wage_rate_from_year aroess case status?

```{r}
# create model
model_wage_rate_from_year <- lm( wage_rate_from_year ~ year, data = LCA_sample)

summary(model_wage_rate_from_year)
```
Visualizing the model predictor


```{r}
LCA_sample %>% 
  ggplot(aes(x = year,
             y = wage_rate_from_year,
             color = case_status)) +
  geom_jitter(alpha = 0.5) +
  geom_smooth(formula = y ~ x, method = "lm") +
  ylim(0, 300000) +
  scale_color_colorblind() 
```
The result shows statistically signification between wage_rate_from_year (the actually 
minimum paid wage) and year. The year has a positive effect on the wage_rate_from_year, 
with a increase of 469055 points on wage_rate_from_year for every point increase on the year. 
The result also shows the US companies need to spend more money on recruiting foreign 
worker and applying for LCA.

## Model 2

> Question: What is the relationship between wage_rate_from_year of denied case_status and state?

```{r}

denied_case <- LCA_application %>% 
  filter(case_status == "denied")

model_prevailing_wage_state <- lm(wage_rate_from_year ~ worksite_state, data = denied_case)

summary(model_prevailing_wage_state)
```
The result (p > 0.05) shows we could not reject null hypothesis, so the denied case wage_rate_from_year have no relationship with state.

## Model 3

> Effect of worksite_state, prevailing_wage_year on wage_rate_from_year of data science related occupations

```{r}
# filter original data set and select data science related jobs.
ds_occupation <- c("machine learning", "data scientist", "data engineer", "data analyst")

LCA_application <- LCA_application %>% 
  mutate(occupation = case_when(grepl("machine learning", job_title, fixed = TRUE) ~ "machine learning",
                                grepl("data scientist", job_title, fixed = TRUE) ~ "data scientist",
                                grepl("data engineer", job_title, fixed = TRUE) ~ "data engineer",
                                grepl("data analyst", job_title, fixed = TRUE) ~ "data analyst",
                                TRUE ~ job_title))

LCA_application %>% 
  filter(occupation %in% ds_occupation) %>% 
  count(occupation) %>% 
  kable()
```

```{r}
LCA_ds <- LCA_application %>% 
  filter(occupation %in% ds_occupation)

model_ds <- lm(wage_rate_from_year ~ prevailing_wage_year + worksite_state + occupation, data = LCA_ds)

summary(model_ds)
```
The result shows only prevailing_wage_year and date scientist occupation has effects on wage_rate_from_year of data science related occupations

```{r}
model_ds <- lm(wage_rate_from_year ~ prevailing_wage_year + occupation, data = LCA_ds)

summary(model_ds)
```
The output shows there is a positive relationship between prevailing_wage_year and wage_rate_from_year of data science related occupations. Data scientist occupation also has a positive relationship.

Let's visualize.

```{r}
LCA_ds %>% 
  ggplot(aes(x = prevailing_wage_year,
             y = wage_rate_from_year,
             color = occupation)) +
  geom_jitter(alpha = .5) +
  facet_wrap(~occupation, scales = "free") +
  geom_smooth(method = "lm") +
  labs(x = "Prevailing Wage Rate",
       y = "Minimum wage rate")
```
```{r}
effect("occupation", model_ds) %>%
  data.frame() %>%
  ggplot(aes(x = reorder(occupation, fit),
             y = fit,
             ymin = lower,
             ymax = upper,
             label = format(fit, digits = 2))) +
  geom_errorbar(width = .4) +
  geom_label() +
  labs(x = "Data Science Occupation",
       y = "Minimum Wage Rate")
```




