---
title: "Final Project"
author: "Quan Gan, Yuxuan chen"
date: "11/27/2021"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(ggthemes)
library(effects)
library(grid)
library(gridExtra)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width=10, fig.height=6)
```
# H-1B LCA petitions analysis from 2016 to 2020

## Introduction
Non-US nationals who wish to work in the United States must first get valid visas.
The H-1B visa is a non-immigrant visa that permits a foreign citizen to work in the United
States in a specialty vocation. This is practically the only approach that international students
could settle in the United States. The US employers will fill petitions on behalf of
the high skilled foreign employees to apply for the H-1B visas. LCA refers to Labor Condition
Application, which is the first process of H-1B applications. LCA is essential
for ensuring that you are paid properly as a foreign worker and that you are not exploited
by US companies. H-1B applications are approved by draw but LCA are approved primarily
by Wage that must be higher than the same level US employees’ minimum wage.
In this project, I’m investigating at what factors impact the approval of an H-1B LCA
application. we wish US companies could reference this research to recruit high skills
workers and help them gain H-1B and also the foreign job seekers or international students
could follow this research to find qualified US companies. The larger questions the
project addresses are how vocations’ needs evolve in the US market. The wage change
and number of applications reflect the market’s needs in the United States. The study
will reveal what factors contribute to the denial of LCA applications and which jobs are
most likely to be approved.The significance of this research is that it may represent the
challenges of acquiring a work permit in the United States as well as trends in the US
market. The project’s societal problems are that it will affect international students’ or
foreign job seekers’ decisions about whether or not to study or settle in the United States. 
Furthermore, the project will affect the US companies about whether or not to hire foreign
employees. We’ll emphasize on data science-related professions like data scientist and
data analyst after analyzing the total LCA applications.

### Research Question
> What factors impact the approval of LCA applications?

> What companies have high expectations of international workers?

> What are the changes of data science jobs; demands in the US market?

## Data 
The data is extracted from non-immigrant and immigrant application tables within
the office of foreign labor certification’s case management systems. The U.S. DEPARTMENT
OF LABOR publishes the disclosure data on their website. https://www.dol.gov/agencies/eta/foreign-labor/performance. 
The data set is LCA Programs (H-1B, H-1B1, E-3). It contains LCA applications xlsx
files from 2016 fiscal year to 2020 fiscal year. For the 2020 fiscal year, there are four xlsx
files so in the later data wrangling, We need to merge them into one table. Every year, the
number of variables grows. There are 41 variables in 2016, but there are 95 variables in
2020. Furthermore, the names of the variables differ. As a result, We’ll use variables that
are present in all fiscal years and have the same meaning.

```{r}
# read data in
LCA_application <- read_csv("Data/LCA_cleaned.csv")
```


### Curation Rationale
The purpose of this data is to investigate the factors that influence the approval of H1B LCA applications, 
the companies that anticipate to recruit more foreign workers, and
the trajectory of data science positions in the US market between 2016 and 2020. Each
original data set contains unique records identifiable by the appropriate OFLC case
number based on the most recent date a case determination judgment was made for the
fiscal year. Furthermore, the data set is massive and imbalanced. The cleaned data set
contains 3004616 records, yet only 1% of applications were declined. The H1-B type is
the only one included in the LCA applications, and wage rates are uniform in year unit.
20% of the entries in the original data set had no full time position values. In order to deal
with those missing values, we analyzed 75% percentile value for part-time positions and
select 87000 as the cut-off. The cleaned columns' names display below.

```{r}
colnames(LCA_application) %>% 
  kable()
```


### Participant Demographic
Because of policies, the public disclosure data set does not contain personally identifiable
information. Therefore, there are no direct demographic information regarding age, gender,
or race and ethnicity information but the data set contains the potential annual wage
and employer name and work site information. We can guarantee that all of the participants are international employees from other nations other than the United States and there
are certain companies in the United States who are interested in hiring them.

### Annotator Demographic
Based on petition information, the data is labeled as certified and denied.. There is no direct demographic characteristics of the annotators but the official policies and regulations
provide guidance about the approval. The U.S. Citizenship and Immigration Services
website (https://www.uscis.gov/working-in-the-united-states/h-1b-specialty-occupations.) 
also demonstrate the LCA requirements.

1. The employer/agent will pay the H-1B worker a wage which is no less than the
wage paid to similarly qualified workers or, if greater, the prevailing wage for the
position in the geographic area in which the H-1B worker will be working.

2. The employer/agent will provide working conditions that will not adversely affect
other similarly employed workers.

3. At the time of the labor condition application there is no strike or lockout at the
place of employment.

4. Notice of the filing of the labor condition application with the DOL has been given to
the union bargaining representative or has been posted at the place of employment.

### Data Collection Situation
Records from the fiscal years 2016 through 2020 are included in the data set. The fiscal
year runs from October 1 of the previous year to September 30 of the current year. For
example, the fiscal year 2020 contains records from October 1, 2019 through September 30, 2020. 
The data is extracted from nonimmigrant and immigrant application tables
within the Office of Foreign Labor Certification’s case management systems. The data is
annotated by the faculties in Office of Foreign Labor certification of Department of Labor.

## Data Visualization

### Visualization One

> Question: What companies have high expectations of international workers arcoss years?

```{r}
# Find the top 10 companies that have most LCA applications.
top_10_company <- LCA_application %>% 
  select(employer_name) %>% 
  add_count(employer_name) %>% 
  arrange(desc(n)) %>% 
  unique() %>% 
  top_n(10)

# Count the LCA applications of top 10 companies across years
company_application_number <- LCA_application %>% 
  filter(employer_name %in% top_10_company$employer_name) %>% 
  select(employer_name, year) %>% 
  group_by(year) %>% 
  add_count(employer_name) %>% 
  arrange(year, desc(n)) %>% 
  unique() 

#Visualizing
company_application_number %>% 
  group_by(employer_name, year) %>% 
  ggplot(aes(x = year,
             y = n,
             fill = year)) +
  geom_col(position = "dodge") + 
  facet_wrap(~employer_name, ncol = 3) +
  coord_flip() +
  labs(y = "Number of LCA applications") +
  scale_color_colorblind() 
  
```

This plot shows the top 10 companies in total that apply for LCA applications from 2016 to 2020.
Each subplot shows their applications trends. We could know for Google LLC and cognizant technology solutions us corp,
they don't have LCA applications in 2016 and 2017. Half of top 10 companies increase their
LCA applications. In addition, I choose the top 10 companies that have most LCA 
applications in total rather than choosing each year top 10 companies because some 
companies might only desire foreign workers in a specific year and analyzing overall top 10 
companies could illustrate the recruiting trends. From this result, we might be surprised that 
the famous big IT companies like Amazon, Google, Microsoft are not the top demand on foreign workers. 
In addition, we want to highlight the institutions. Some companies changed their name over years like 
Amazon but the USCIS would no update the new name to the old records.

### Visualization Two

> Question: Would the prevailing wage rate and paid wage rate influence the approval of LCA application?

In LCA, the prevailing wage is defined as "the average wage paid to similarly employed
workers in the requested occupation in the area of intended employment" from 
https://www.usavisanow.com/h-1b-visa/h1b-visa-resources/prevailing-wage/.

The paid wage rate is the actually wage paid to the foreign employee. 


```{r}
# count case_status
LCA_application %>% count(case_status) %>% 
  kable()
```

In this data set, there are over 300 million records but only 1% applications failed,
which is extremely unbalanced. In order to explore the correlation among prevailing 
wage rate, paid wage rate, and the approval of LCA application, we decide to sample
same amount record from data set.

```{r}

# set seed
set.seed(1)

# sample records from dataset
certified <- LCA_application %>% 
  filter(case_status == "certified") %>% 
  sample_n(10000)

denied <- LCA_application %>% 
  filter(case_status == "denied") %>% 
  sample_n(10000)


LCA_sample <- rbind(certified, denied)

# The wage_rate_from_year is the minimum paid wage rate and wage_rate_to_year is the
# maximum paid wage rate. So, we need to calculate the mean wage rate.
LCA_sample <- LCA_sample %>% 
  mutate(mean_wage_rate_year = ifelse(wage_rate_to_year == 0, 
                                      wage_rate_from_year,
                                      (wage_rate_from_year + wage_rate_to_year)/2))

LCA_sample <- LCA_sample %>% 
  mutate(difference_prevailing_wage = mean_wage_rate_year - prevailing_wage_year)

```

```{r}
LCA_sample %>% 
  group_by(case_status) %>% 
  ggplot(aes(x = factor(year),
             y = prevailing_wage_year,
             fill = case_status)) +
  geom_boxplot() +
  coord_cartesian(ylim=c(0,150000)) +
  labs(x = "Year",
       y = "Prevailing wage rate unit year",
       fill = "Case Status") +
  scale_color_colorblind() -> prevailing_plot_1

LCA_sample %>% 
  group_by(case_status) %>% 
  ggplot(aes(x = factor(year),
             y = mean_wage_rate_year,
             fill = case_status)) +
  geom_boxplot() +
  coord_cartesian(ylim=c(0,150000))+
  labs(x = "Year",
       y = "Paid wage rate unit year",
       fill = "Case Status") -> mean_wage_plot_1

LCA_sample %>% 
  group_by(case_status) %>% 
  ggplot(aes(x = factor(year),
             y = difference_prevailing_wage,
             fill = case_status)) +
  geom_boxplot() +
  coord_cartesian(ylim=c(-10000,25000))+
  labs(x = "Year",
       y = "Difference of prevailing wage rate and paid wage rate",
       fill = "Case Status") -> different_prevailing_plot_1

grid.arrange(prevailing_plot_1, mean_wage_plot_1, different_prevailing_plot_1, ncol = 2)
```

The means of prevailing wage rate and paid wage rate of refused LCA applications 
are lower than the means of certified LCA applications, as seen in the first two box plots. 
The mean gap between paid and prevailing pay rates is greater than the mean difference 
between paid and prevailing wage rates.
The third box plots shows that the applicants receive the denied results because 
the employers don't pay enough wage rate comparing to the prevailing wage rate.
In addition, some paid wage rates of denied applications are less than the 
prevailing wage rate but certified applications always provide wage rate larger 
than prevailing wage. In the data, there are some extreme high wage rates, which 
are noise. In order to help the users understand the visualization better, I change 
the y-axis so the audiences could focus on the mean values.

### Visualization Three

> Question: What is the trends of data science related jobs?

In this question, we will analyze four popular data science related jobs, machine learning,
data scientist, data engineer, and data analyst. How would 
their demand change? How would their wage rate change?

```{r}
# filter original data set and select data science related jobs.
ds_occupation <- c("machine learning", "data scientist", "data engineer", "data analyst")

LCA_application <- LCA_application %>% 
  mutate(occupation = case_when(grepl("machine learning", job_title, fixed = TRUE) ~ "machine learning",
                                grepl("data scientist", job_title, fixed = TRUE) ~ "data scientist",
                                grepl("data engineer", job_title, fixed = TRUE) ~ "data engineer",
                                grepl("data analyst", job_title, fixed = TRUE) ~ "data analyst",
                                TRUE ~ job_title))

LCA_application %>% 
  filter(occupation %in% ds_occupation) %>% 
  count(occupation) %>% 
  kable()
```
```{r}
# Visualizing LCA applications amount of data science jobs.
LCA_application %>% 
  filter(occupation %in% ds_occupation) %>% 
  group_by(occupation) %>% 
  ggplot(aes(x = year, fill = factor(year))) +
  geom_bar(stat="count", position = "dodge") +
  geom_label(stat='count', aes(label=..count..), hjust = -0.1) +
  scale_y_continuous(limits=c(0, 6000)) +
  scale_color_colorblind() +
  coord_flip() +
  facet_wrap(~occupation) + 
  labs(x = "Year",
       y = "Number of Applications",
       fill = "Year") 
```
This bar chart depicts the number of LCA applications. of data science-related 
professions over time. With each passing year, the need for data science-related 
occupations grows. Particularly, the data engineer occupation growth fast. 
In addition, to assist viewers comprehend the graphic, I indicate the specific 
number of LCA applications. Despite the fact that data science is becoming more 
popular, it still still accounts for a small portion of demand in the US market, 
which totals 300 million applications.

```{r}
# Visualizing the prevailing wage rage of data science job
LCA_application %>% 
  filter(occupation %in% ds_occupation) %>% 
  group_by(occupation, year) %>% 
  summarise(mean_wage = mean(prevailing_wage_year)) %>% 
  ggplot(aes(x = year, y = mean_wage, color = occupation)) +
  geom_line(size = 1) +
  labs(x = "Year",
       y = "Mean of Prevailing wage Rate",
       fill = "Occupation") -> mean_prevailing_plot

# Visualizing the mean wage rage of data science job
LCA_application %>% 
  filter(occupation %in% ds_occupation) %>% 
  group_by(occupation, year) %>% 
  summarise(mean_wage = mean(wage_rate_from_year)) %>% 
  ggplot(aes(x = year, y = mean_wage, color = occupation)) +
  geom_line(size = 1) +
  labs(x = "Year",
       y = "Mean of Minimum paid wage Rate",
      fill = "Occupation") -> mean_minimum_paid_plot

grid.arrange(mean_prevailing_plot, mean_minimum_paid_plot, ncol = 1)
```

The two graphs above show pay trends for data science positions. Except for the 
data engineer occupation, which has seen a significant increase in wage rates in 2018, 
other data science jobs have seen just minor changes. They raise the paid minimum 
wage rate for those who seek to recruit overseas labor. The data scientist occupation 
has the quickest income increase and is also one of the most in-demand positions 
in the data science field.We use different color lines to represent different jobs 
and alter the line width to assist users comprehend the graphic.

## Data Modeling

### Model 1

> Question: What is the relationship of year and wage_rate_from_year aroess case status?

```{r}
# create model
model_wage_rate_from_year <- lm( wage_rate_from_year ~ year, data = LCA_sample)

summary(model_wage_rate_from_year)
```

Visualizing the model predictor


```{r}
LCA_sample %>% 
  ggplot(aes(x = year,
             y = wage_rate_from_year,
             color = case_status)) +
  geom_jitter(alpha = 0.5) +
  geom_smooth(formula = y ~ x, method = "lm") +
  ylim(0, 300000) +
  scale_color_colorblind() 
```

The result shows statistically signification between wage_rate_from_year (the actually 
minimum paid wage) and year. The year has a positive effect on the wage_rate_from_year, 
with a increase of 469055 points on wage_rate_from_year for every point increase on the year. 
The result also shows the US companies need to spend more money on recruiting foreign 
worker and applying for LCA.

### Model 2

> Question: What is the relationship between wage_rate_from_year of denied case_status and state?

```{r}

denied_case <- LCA_application %>% 
  filter(case_status == "denied")

model_prevailing_wage_state <- lm(wage_rate_from_year ~ worksite_state, data = denied_case)

summary(model_prevailing_wage_state)
```

The result (p > 0.05) shows we could not reject null hypothesis, so the denied case wage_rate_from_year have no relationship with state.

### Model 3

> Effect of worksite_state, prevailing_wage_year on wage_rate_from_year of data science related occupations

```{r}
LCA_ds <- LCA_application %>% 
  filter(occupation %in% ds_occupation)

model_ds <- lm(wage_rate_from_year ~ prevailing_wage_year + worksite_state + occupation, data = LCA_ds)

summary(model_ds)
```
The result shows only prevailing_wage_year and date scientist occupation has effects on wage_rate_from_year of data science related occupations

```{r}
model_ds <- lm(wage_rate_from_year ~ prevailing_wage_year + occupation, data = LCA_ds)

summary(model_ds)
```
The output shows there is a positive relationship between prevailing_wage_year and wage_rate_from_year of data science related occupations. Data scientist occupation also has a positive relationship.

Let's visualize.
```{r}
LCA_ds %>% 
  ggplot(aes(x = prevailing_wage_year,
             y = wage_rate_from_year,
             color = occupation)) +
  geom_jitter(alpha = .5) +
  facet_wrap(~occupation, scales = "free") +
  geom_smooth(method = "lm") +
  labs(x = "Prevailing Wage Rate",
       y = "Minimum wage rate")
```
```{r}
effect("occupation", model_ds) %>%
  data.frame() %>%
  ggplot(aes(x = reorder(occupation, fit),
             y = fit,
             ymin = lower,
             ymax = upper,
             label = format(fit, digits = 2))) +
  geom_errorbar(width = .4) +
  geom_label() +
  labs(x = "Data Science Occupation",
       y = "Minimum Wage Rate")
```

As we can see above, the data scientist and machine learning positions' minimum paid wages keep increasing as the 
increment of the prevailing wage but the data analyst and data engineer don't. In addition, the machine learning 
position has the most variances. The possible reasons for this results is the USCIS doesn't consider the position levels 
as a factor in the LCA applications. Even for the same position title, the wage rates are different based on the 
positions levels. If we sum the data science related job amount, we would notice they are just a tiny port of LCA applications, 50k out of 3M. In another words, the US companies are not likely to hire foreign workers for Data science positions. 